Symbolic music generation using Mamba architecture

### 2024-06-13 docs/paper/paper.tex
add mamba section

\addbibresource{bibliografia.bib}

@@

\section{Architecture}
We decided to use the current state of the art SSM architecture called Mamba \cite{mamba}. It further improves the Structured State Space sequence (S4) \cite{s4} model by allowing the model to selectively propagate or forget information along the
sequence length dimension depending on the current token. It also introduced an efficient hardware-aware parallel algorithm to speed up the computation time. Mamba achieves 5× higher throughput than Transformers and linear scaling in sequence length \cite{mamba}. It reaches state-of-the-art performance across several modalities such as language, audio, and genomics. On language modeling, our Mamba-3B model
outperforms Transformers of the same size and matches Transformers twice its size, both in pretraining and downstream evaluation \cite{mamba}.

>>>>

\addbibresource{references.bib}

@@

\section{Architecture}
Mamba utilises a novel approach called selective state space models (SSMs), which enhance traditional state space models by allowing 
parameters to be functions of the input. This design enables the model to selectively propagate or forget information along the sequence length,
dependent on the current input token. Mamba integrates this selective SSM approach into a simplified end-to-end architecture, foregoing traditional components 
like attention or MLP blocks found in architectures like Transformers.

The main difference between Selective SSM and traditional SSM is the input-dependence, which mimics what the attention mechanism in Transformers
does—essentially assessing whether a specific token is important or not. However, this feature sacrifices the ability of traditional SSMs
to precompute all learnable matrices ($\Delta$, A, B, C) and their configurations based on sequence length in a single inference pass. 
To address this, we introduce a mechanism of Parallel Associative Scan (similar to Prefix Sum) that requires storing precomputed calculations,
leading to higher memory usage but still maintaining linear computation.

\begin{figure}[!htbp]
    \centering \includegraphics[width=\linewidth]{../assets/SelectiveSSM.png}
    \caption{Selective SSM \cite{mamba}}
    \label{fig:ssm}
\end{figure}

To enhance efficiency further, the authors proposed using a hardware-aware algorithm that leverages two main types of GPU memory: 
SRAM, which is fast but has limited capacity (ideal for matrix calculations), and HBM, which is slower but has a larger capacity. 
The main bottleneck of this approach is managing the data transfer between these memory types.

\begin{figure}[!htbp]
    \centering \includegraphics[width=\linewidth]{../assets/Mamba.png}
    \caption{Mamba block \cite{mamba}}
    \label{fig:mamba}
\end{figure}

Selective SSM is a crucial component of the Mamba block, but the system also includes linear layers that increase dimensionality, nonlinear layers, and gating connections.
Whole architecture is build from many Mamba block which are computed layer by layer.

\begin{table}[t] \centering
    % Podpis tabeli umieszczamy od góry
    \caption{Model comparison \cite{mamba}}
    \label{tab:comparison}

    % Tabela z trzema kolumnami:
    % dwie wyrównanie do środka [c], a ostatnia do prawej [r]
    % szerokość kolumn automatyczna (równa szerokości tekstu)
    \resizebox{\linewidth}{!}{\begin{tabular}{| c | c | c | c |} \hline
        Architecture       		& Complexity    			& Memory   		& Performance \\ \hline\hline
        Transformer      			&    	 $ O(N^2)$ 		& $O(N^2)$     		& great    \\ \hline
        RNN     				&    	$O(N)$		 	& $O(N)$    		& poor     \\ \hline
        SSM      				&  	$O(N)$ 			& $O(N)$     		& poor     \\ \hline
       Selective SSM (Mamba)     &    	 $O(N)$ 			& $O(N)$    		& great.  \\ \hline
        % Komórka o szerokości dwóch kolumn, wyrównana do prawej
        % Przypisy dolne w tabelach wstawiamy przez \tablefootnote
        % \multicolumn{2}{|r|}{Suma\tablefootnote{Table footnote.}} & 123,45 \\ \hline
    \end{tabular}}
\end{table}

\begin{figure}[!htbp]
    \centering \includegraphics[width=\linewidth]{../assets/results.png}
    \caption{Benchmarks \cite{mamba}}
    \label{fig:mamba}
\end{figure}
